---
title: "Week 4 Assignment"
author: "David Russo"
date: "2/15/2017"
output: word_document
---

* 4.7 #4
```{r set_up_libraries, echo = FALSE, include = FALSE}
library(tidyverse)
library(MASS)
```

  + a. 
In this scenario, 10% of the observations will be used to make a prediction.
  
  + b.
In this scenario, 1% of the observations will be used to make a prediction.
  
  + c.
In this scenario, $10^{-100}$ of the observations will be used to make a prediction. 

  + d.
Consider a scenario where there are n = 1000 observations. If p = 1, we will utilize 100 observations per prediction on average. If p = 2, we will use 10 observations per prediction on average. If p = 100, we will use approximately 0 observations per prediction on average. In order to use even 10 observations on average, we would need n = $10^21$ observations when p = 100. The plot below demonstrates the percentage of observations used as a function of p (note that the y-axis is on a log scale).

```{r, echo = FALSE}

dat <- data.frame(
  p = seq(from = 1, to = 10, by = 1),
  proportion = sapply(c(1:10), function(x) 10^-x)
)

dat %>% 
  ggplot(aes(x = p, y = proportion)) +
    geom_line() + 
      scale_y_log10() + 
        scale_x_continuous(breaks = 1:10) + 
          ylab("Proportion of Observations Used \n Log Scale")

```



  + e.
When p = 1, the hypercube (in this case, a 1D line) will have length 0.10. 
When p = 2, the hypercube (in this case, a 2D square) will have length $\sqrt{0.10}$ = 0.316. 
When p = 100, the hypercube (in this case, a 100D cube) will have length $0.10^{1/100}$ = 0.977.
In general, for a given p, the hypercube will require a length of $0.10^{1/p}$. For an arbitrary percentage of the training observations $t$ for a hypercube of dimension $p$, the hypercube will need to be of length $t^{1/p}$. 
