---
title: "Week 9 Assignment"
author: "David Russo"
date: "3/12/2017"
output: pdf_document
---

```{r overhead, echo = TRUE, include = FALSE}
library(tidyverse)
library(ISLR)
```

10.7 #7
```{r demonstration_strange_wording, echo = TRUE, include = TRUE}
# load the data
data("USArrests")

# scale the data so that each variable has a mean of 0 and a variance of 1
usa_scaled <- as.data.frame(scale(USArrests))
# verify that the scaled data's variables all have mean 0 and variance 1
apply(usa_scaled, 2, function(x) round(c(mean(x), var(x)), 2))

# create data frame of combinations of rows of usa_scaled data set
# there are choose(50, 2) = 1225 ways to choose 2 observations out of 50
combs <- t(combn(nrow(usa_scaled), 2))
head(combs)

###----------------- Determine Correlations ----------------------------

# cycle through combs to get each correlation between rows
one_minus_cors <- rep(0, nrow(combs))

# create function for getting correlations between each observation
one_minus_cor_usa <- function(x){
  res <- cor(as.numeric(usa_scaled[combs[x, 1], ]),
             as.numeric(usa_scaled[combs[x, 2], ]))
  res <- 1 - res
  res
}

# determine 1 - r_ij for each pair of observations
one_minus_cors <- sapply(1:nrow(combs), function(x) one_minus_cor_usa(x))

###----------------- Determine Squared Euclidean Distances --------------

# cycle through combs to get distance between rows
distances <- dist(usa_scaled)^2

###----------------- Determine Proportionality -------------------
summary(one_minus_cors/distances)

hist(one_minus_cors/distances)

```

As can be seen from the summary above, there is no evidence of proportionality. If there were proportionality, there would only be one value. I believe the wording of this question is either wrong or simply hard to understand. The question refers to distances and correlations between the $i^{th}$ and $j^{th}$ observations, which would lead me to believe it is referring to the rows of the data set (i.e., 50 rows; one per state). If the authors meant to refer to the distances and correlations between $i^{th}$ and $j^{th}$ variables (i.e., the 4 variables of Murder, Assault, UrbanPop, and Rape), then there is constant proportionality. This is shown below.

```{r possible_wording_meaning}
# squared distance matrix 
distance_matrix <- dist(t(usa_scaled))^2
distance_matrix

# 1 - correlation matrix
one_minus_cor_mat <- as.dist(1-cor(usa_scaled))
one_minus_cor_mat

# demonstration of proportionality 
distance_matrix/one_minus_cor_mat
```

10.7 #10

  * a)
```{r 10a, echo = TRUE, include = TRUE}
# generate raw data matrix 
set.seed(2017)
dat <- matrix(rnorm(60*50) + rep(c(2, -5, 7), each = 20), ncol = 50)
true_labels <- rep(c("group 1", "group 2", "group 3"), each = 20)

# get means for first 20 rows, representing the first class
mean(dat[1:20, ])

# get means for the second 20 rows, representing the second class
mean(dat[21:40, ])

# get means for the final 20 rows, representing the third class
mean(dat[41:60, ])
```

  * b)
```{r 10b, echo = TRUE, include = TRUE}
# perform principal components
pr_10b <- prcomp(dat, scale = TRUE)

# plot the first two principal components
plot_dat <- as.data.frame(pr_10b$x[, c(1, 2)])
plot_dat$label <- true_labels

plot_dat %>%
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(col = true_labels)) + 
  xlab("Principal Component 1") + 
  ylab("Principal Component 2")
```

  * c)
```{r 10c, echo = TRUE, include = TRUE}
km_10c <- kmeans(dat, 3, nstart = 20)
dat_10c <- as.data.frame(dat)
dat_10c$true_label <- true_labels
dat_10c$km_label <- km_10c$cluster

table(dat_10c$true_label, dat_10c$km_label)
```

  * d)
```{r 10d, echo = TRUE, include = TRUE}
km_10d <- kmeans(dat, 2, nstart = 20)
dat_10d <- as.data.frame(dat)
dat_10d$true_label <- true_labels
dat_10d$km_label <- km_10d$cluster

table(dat_10d$true_label, dat_10d$km_label)
```



