---
title: "Week 8 Assignment"
author: "David Russo"
date: "3/9/2017"
output: pdf_document
---

```{r overhead, echo = FALSE, include = FALSE}
library(tidyverse)
library(ISLR)
```

10.7 #3

  * a)
```{r a, echo = TRUE, include = TRUE}
obs <- data.frame(
  X1 = c(1, 1, 0, 5, 6, 4),
  X2 = c(4, 3, 4, 1, 2, 0)
)

obs %>%
  ggplot(aes(x = X1, y = X2)) +
  geom_point()

```

  * b)
```{r b, echo = TRUE, include = TRUE}
# set random number seed for replicable results
set.seed(2017)

# assign the labels
obs$cluster <- sample(rep(c(1, 2), each = 3), 6, replace = FALSE)

# print the labels
obs

# plot the labels
obs %>%
  dplyr::select(X1, X2, cluster) %>% 
  ggplot(aes(x = X1, y = X2, color = as.factor(cluster))) +
  geom_point()
```

  * c)
```{r c, echo = TRUE, include = TRUE}
centroids <-
  obs %>%
  dplyr::group_by(cluster) %>%
  dplyr::summarise(centroid_X1 = round(mean(X1), 2), centroid_X2 = round(mean(X2), 2)) %>%
  as.data.frame()

centroids
  
```

  * d)
```{r d, echo = TRUE, include = TRUE}

# create distance from cluster1 centroid
obs$dist_from_cluster1_centroid <- round(
  sqrt((obs$X1 - centroids$centroid_X1[centroids$cluster == 1])^2 + 
       (obs$X2 - centroids$centroid_X2[centroids$cluster == 1])^2), 4)

# create distance from cluster2 centroid
obs$dist_from_cluster2_centroid <- round(
  sqrt((obs$X1 - centroids$centroid_X1[centroids$cluster == 2])^2 + 
       (obs$X2 - centroids$centroid_X2[centroids$cluster == 2])^2), 4)

obs$new_cluster <-
  ifelse(obs$dist_from_cluster1_centroid <= obs$dist_from_cluster2_centroid, 1, 2)

obs

```

  * e)
```{r e, echo = TRUE, include = TRUE}

cluster_difference <- FALSE

while(cluster_difference == FALSE){

# reset obs$cluster to obs$new_cluster
obs$cluster <- obs$new_cluster
  
# create distance from cluster1 centroid
obs$dist_from_cluster1_centroid <- round(
  sqrt((obs$X1 - centroids$centroid_X1[centroids$cluster == 1])^2 + 
       (obs$X2 - centroids$centroid_X2[centroids$cluster == 1])^2), 4)

# create distance from cluster2 centroid
obs$dist_from_cluster2_centroid <- round(
  sqrt((obs$X1 - centroids$centroid_X1[centroids$cluster == 2])^2 + 
       (obs$X2 - centroids$centroid_X2[centroids$cluster == 2])^2), 4)

obs$new_cluster <-
  ifelse(obs$dist_from_cluster1_centroid <= obs$dist_from_cluster2_centroid, 1, 2)

cluster_difference <- all(obs$cluster == obs$new_cluster)

print(obs)

}

```

The clustering only took one iteration until the algorithm converged.

  * f)
```{r f, echo = TRUE, include = TRUE}
obs %>%
  dplyr::select(X1, X2, new_cluster) %>% 
  ggplot(aes(x = X1, y = X2, color = as.factor(new_cluster))) +
  geom_point()
```

10.7 #7
```{r demonstration_strange_wording, echo = TRUE, include = TRUE}
# load the data
data("USArrests")

# scale the data so that each variable has a mean of 0 and a variance of 1
usa_scaled <- as.data.frame(scale(USArrests))
# verify that the scaled data's variables all have mean 0 and variance 1
apply(usa_scaled, 2, function(x) round(c(mean(x), var(x)), 2))

# create data frame of combinations of rows of usa_scaled data set
# there are choose(50, 2) = 1225 ways to choose 2 observations out of 50
combs <- t(combn(nrow(usa_scaled), 2))
head(combs)

###----------------- Determine Correlations ----------------------------

# cycle through combs to get each correlation between rows
one_minus_cors <- rep(0, nrow(combs))

# create function for getting correlations between each observation
one_minus_cor_usa <- function(x){
  res <- cor(as.numeric(usa_scaled[combs[x, 1], ]),
             as.numeric(usa_scaled[combs[x, 2], ]))
  res <- 1 - res
  res
}

# determine 1 - r_ij for each pair of observations
one_minus_cors <- sapply(1:nrow(combs), function(x) one_minus_cor_usa(x))

###----------------- Determine Squared Euclidean Distances --------------

# cycle through combs to get distance between rows
distances <- dist(usa_scaled)^2

###----------------- Determine Proportionality -------------------
summary(one_minus_cors/distances)

hist(one_minus_cors/distances)

```

As can be seen from the summary above, there is no evidence of proportionality. If there were proportionality, there would only be one value. I believe the wording of this question is either wrong or simply hard to understand. The question refers to distances and correlations between the $i^{th}$ and $j^{th}$ observations, which would lead me to believe it is referring to the rows of the data set (i.e., 50 rows; one per state). If the authors meant to refer to the distances and correlations between $i^{th}$ and $j^{th}$ variables (i.e., the 4 variables of Murder, Assault, UrbanPop, and Rape), then there is constant proportionality. This is shown below.

```{r possible_wording_meaning}
# squared distance matrix 
distance_matrix <- dist(t(usa_scaled))^2
distance_matrix

# 1 - correlation matrix
one_minus_cor_mat <- as.dist(1-cor(usa_scaled))
one_minus_cor_mat

# demonstration of proportionality 
distance_matrix/one_minus_cor_mat




```





